All the concepts behind TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages (Draft)

### 2020-11-06 article.tex
Articulo gaming: un poquito de modificaci'on en la intro

\en{Understanding the factors impacting in the positive or negative changes of skill could be used to develop new training strategies or adapting the environment to avoid unnecessary tasks.}
\es{Conocer c\'omo cambian la habilidades de la personas en el tiempo es crucial en muchas \'areas.}
%
\en{For example, in sports development and education, knowing the skill and how it changes can support avoiding exposing agents to imbalanced challenges, maintaining the interest and engagement, and monitoring the learning and improvement process.}
\es{En contextos educativos, calificaciones precisas pueden servir para evitar exponer a los agentes a desaf\'ios desproporcionados, estimulando as\'i el inter\'es y la participaci\'on.}
%
\en{In practice, since skill is a hidden variable, the best we can do is estimating it from its direct observable consequences.}
\es{En la pr\'actica, dado que la habilidad es una variables oculta, lo mejor que podemos hacer es estimarla a partir de sus consecuencias observables m\'as directas.}
%
\en{The most widely used sources of information in education and sports are the outcomes of problem solving and competition between opponents.}
\es{Las fuentes de informaci\'on m\'as utilizadas en educaci\'on y deportes provienen del producto de resoluci\'on de problemas y de competencia entre oponentes.}
%
\en{The naive solution (i.e. frequency of positive results) is not a good estimator because it depends on the difficulty of the challenges.}
\es{La soluci\'on ingenua, frecuencia de resultados positivos, no es un buen indicador fundamentalmente porque depende tambi\'en de la dificultad de los desaf\'ios.}
%
\en{The estimation should rely on pairwise comparisons.}
\es{Por esa raz\'on, todos las soluciones se basan en comparaciones entre pares.}
%
\en{Since the first generative models, proposed almost a century ago by~\cite{thurstone1927-comparativeJugement} and~\cite{zermelo1929-translation}, it is assumed that the probability of an observed result, $r_{ij}$, depends on the difference in performance, $d_{ij}$, between the agent $i$ and target $j$, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
\es{Ya los primeros modelos generativos, propuestos hace casi un siglo por~\cite{thurstone1927-comparativeJugement} y~\cite{zermelo1929-translation}, suponen que la probabilidad de un resultado observado, $r$, depende de la diferencia de rendimiento, $d$, entre el agente y su objetivo, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
%
\en{The field revitalized 20 years later by the publications of~\cite{bradley1952-pairedComparisions} and~\cite{mosteller1951-remarks}.}
\es{El campo revitaliz\'o 20 a\~nos despu\'es gracias a las publicaciones de \cite{bradley1952-pairedComparisions} y~\cite{mosteller1951-remarks}.}

>>>>

\en{Understanding the factors impacting in the positive or negative changes of skill could be used to develop new training strategies or adapting the environment to avoid unnecessary tasks.}
\es{Conocer c\'omo cambian la habilidades de la personas en el tiempo es crucial en muchas \'areas.}
%
\en{For example, in sports development and education, knowing the skill and how it changes can support avoiding the exposition of agents to imbalanced challenges, maintaining the interest and engagement, and monitoring the learning and improvement process.}
\es{En contextos educativos, calificaciones precisas pueden servir para evitar exponer a los agentes a desaf\'ios desproporcionados, estimulando as\'i el inter\'es y la participaci\'on.}
%
\en{In practice, since skill is a hidden variable, the best we can do is estimating it from its direct observable consequences.}
\es{En la pr\'actica, dado que la habilidad es una variables oculta, lo mejor que podemos hacer es estimarla a partir de sus consecuencias observables m\'as directas.}
%
\en{The most widely used information sources in education and sports are the outcomes of problem solving and competition results.}
\es{Las fuentes de informaci\'on m\'as utilizadas en educaci\'on y deportes provienen del producto de resoluci\'on de problemas y de competencia entre oponentes.}
%
\en{The naive solution (i.e. frequency of winning or passing the exam) is not a good estimator because it depends on the difficulty of the challenges.}
\es{La soluci\'on ingenua, frecuencia de resultados positivos, no es un buen indicador fundamentalmente porque depende tambi\'en de la dificultad de los desaf\'ios.}
%
\en{The estimation should rely on pairwise comparisons.}
\es{Por esa raz\'on, todos las soluciones se basan en comparaciones entre pares.}
%
\en{Since the first generative models, proposed almost a century ago by~\cite{thurstone1927-comparativeJugement} and~\cite{zermelo1929-translation}, it is assumed that the probability of an observed result, $r_{ij}$, depends on the difference in performance, $d_{ij}$, between the agent $i$ and target $j$, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
\es{Ya los primeros modelos generativos, propuestos hace casi un siglo por~\cite{thurstone1927-comparativeJugement} y~\cite{zermelo1929-translation}, suponen que la probabilidad de un resultado observado, $r$, depende de la diferencia de rendimiento, $d$, entre el agente y su objetivo, $P(\, r_{ij} \,|\, d_{ij} \,)$.}
%
\en{The field revitalized $20$ years later by the publications of~\cite{bradley1952-pairedComparisions} and~\cite{mosteller1951-remarks}.}
\es{El campo revitaliz\'o 20 a\~nos despu\'es gracias a las publicaciones de \cite{bradley1952-pairedComparisions} y~\cite{mosteller1951-remarks}.}
